### 2014-03-28 brains-feature-cards.tex
Move figures around, and create one page of figures per feature

\clearpage
\onecolumn
\appendix
\section{Feature figures}

% aoa_Kuperman ===============================
\subsection{aoa Kuperman}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/aoa_Kuperman/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/aoa_Kuperman/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/aoa_Kuperman/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/aoa_Kuperman/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% cmu_MNphonemes ===============================
\clearpage
\subsection{cmu MNphonemes}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNphonemes/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/cmu_MNphonemes/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNphonemes/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/cmu_MNphonemes/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% cmu_MNsyllables ===============================
\clearpage
\subsection{cmu MNsyllables}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNsyllables/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/cmu_MNsyllables/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/cmu_MNsyllables/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/cmu_MNsyllables/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_BCs ===============================
\clearpage
\subsection{fa BCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_BCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_BCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_BCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_BCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_CCs ===============================
\clearpage
\subsection{fa CCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_CCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_CCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_CCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_CCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_degrees ===============================
\clearpage
\subsection{fa degrees}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_degrees/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_degrees/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_degrees/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_degrees/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% fa_PR_scores ===============================
\clearpage
\subsection{fa PR scores}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/fa_PR_scores/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/fa_PR_scores/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/fa_PR_scores/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/fa_PR_scores/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% mt_frequencies ===============================
\clearpage
\subsection{mt frequencies}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/mt_frequencies/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/mt_frequencies/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/mt_frequencies/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/mt_frequencies/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% mt_start_frequencies ===============================
\clearpage
\subsection{mt start frequencies}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/mt_start_frequencies/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/mt_start_frequencies/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/mt_start_frequencies/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/mt_start_frequencies/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_BCs ===============================
\clearpage
\subsection{wn BCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_BCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_BCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_BCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_BCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_CCs ===============================
\clearpage
\subsection{wn CCs}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_CCs/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_CCs/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_CCs/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_CCs/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_degrees ===============================
\clearpage
\subsection{wn degrees}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_degrees/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_degrees/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_degrees/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_degrees/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_MNSyns ===============================
\clearpage
\subsection{wn MNSyns}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_MNSyns/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_MNSyns/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_MNSyns/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_MNSyns/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_NSigns ===============================
\clearpage
\subsection{wn NSigns}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_NSigns/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_NSigns/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_NSigns/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_NSigns/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}

% wn_PR_scores ===============================
\clearpage
\subsection{wn PR scores}
\begin{figure*}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{images/feature-cards/wn_PR_scores/feature_variations.png}
    \caption{Feature variation on substitution}
	\subfloat[][Scatter plot]{
		\includegraphics[width=0.3\textwidth]{images/feature-cards/wn_PR_scores/susceptibility_scatter.png}
	}
	~
	\subfloat[][Bin averages and KDE]{
		\includegraphics[width=0.6\textwidth]{images/feature-cards/wn_PR_scores/susceptibility.png}
	}
	\caption{Substitution susceptibility}
    \includegraphics[width=\textwidth]{images/feature-cards/wn_PR_scores/timebags_evolution-hist.png}
	\caption{Timebag evolution}
\end{figure*}