### 2016-07-28 brains-appendix.tex
feat/results: finish presentation of results

\section{Sentence-relative feature effects}

\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/paper-susceptibilities-sentencequantiles.png}
    \caption{\todo{Bring back to 1, and use log-y}}
    \label{fig:feature-susceptibilities-in_sentence}
\end{figure*}